Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-23 cav2017/cav2017.tex
more progress on the AA section (not done though)

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Verification of a controllable system}
\label{sec:cof_verification}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

Let us recall the formulas used for single-input and single-output (SISO) systems 
in sections \ref{sec:reachable} and \ref{sec:observable}.
A controllable system with quantization will have the following dynamics:
\todo{eliminate vectors}
%
\begin{align}
\label{eq:observer_LTI_cf}
\vec{x}_{k+1}=&\mat{A}_{t}\vec{x}_k+\mat{B}_{tp} \vec{r}_k+\mat{B}_{tn} (\vec{\nu}_1 + \vec{\nu}_2 + \vec{\nu}_3)
\end{align}
\begin{align*}
\mat{A}_{t}=&\mat{T}\mat{A}_d\mat{T}^{-1}&
\mat{B}_{tp}=&\mat{T}\mat{B}_d&
\mat{B}_{tn}=&\mat{T}\mat{B}_d&
\mat{T}=&\mat{R}_{cf}\mat{R}^{-1}&
\end{align*}

@@

For a stabilised closed-loop, which is necessary for our safety property, we need
these roots to remain within the unit circle, which we verify using Jury's criteria.
Notice that in our polynomials $c_0=1$ and we only use reals.
\todo{where is $\phi_{input}$ computed?}
Once we have established convergence, we may examine the 
specification of the state and input spaces ($\phi_{safety} \wedge \phi_{input}$).
% Whilst there are other properties, we may include in the verification, 
% we will solely focus on these for simplicity of the explanation.
% We assume that we do not want saturation in our system, 
% hence the control loop has only one mode.

We have also identified 
the quantization noise as a bounded non-deterministic input.
%
The first step constraints
\begin{equation}
\mat{K} \models \vec{u}_k=-\mat{K}\vec{x}_k : \phi_{safety} \wedge \phi_{input}
\label{eq:of_bounds}
\end{equation}
The second step is evaluating the progression of the system over time to evaluate 
$\vec{x}_{k+1} \models \phi_{safety}$. For this step we require 
an initial set from which the system progresses. Obviously, this set $\vec{x}_0$ 
must satisfy the specification; otherwise, the system will be unsafe to begin with.
We accept a specification in the form $\mat{E}\mat{T}\vec{x}_0<\mat{f}$. 
The presence of $\mat{T}$ is because this will typically be given in the original 
state-space. Accelerating \eqref{eq:observer_LTI_cf}, we obtain:
%
\begin{align}
\label{eq:acc_observer_LTI_cf}
\vec{x}=&\mat{A}_t^k\vec{x}_0
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tp} \vec{r}_i
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tn}(\vec{\nu}_1+\vec{\nu}_2+\vec{\nu}_3) : \vec{x}_0 \models \phi_{init}
\end{align}
which would still require us to verify the system for every $k$ up to infinity. 
Instead we use abstract acceleration~\cite{cattaruzza2015unbounded}, which translates the formula into
%
\begin{align}
\label{eq:aa_observer_LTI_cf}
\hat{X}^\#
=&\mathcal{A}_t X_0+\mathcal{B}_{tp} R + \mathcal{B}_{tn} N\\
X_0 =&\left \{\vec{x} : \vec{x} \models \phi_{init} \right\}\nonumber\\
 R =&\{r : \mat{E}_r < \vec{f}_r \}\nonumber\\
 N=&\left \{ \vec{\nu}_1+\vec{\nu}_2+ \vec{\nu}_3 : \nu_1 \in \left[-\frac{q1}{2}\ \ \frac{q1}{2}\right] \wedge \nu_2 \in \left[-\frac{q2}{2}\ \ \frac{q2}{2}\right]  \wedge \nu_3 \in \left[-q3\ \ q3\right]  \right \}\nonumber
\end{align}
%
where $\mathcal{A}_t=\bigcup_{k=1}^\infty \mat{A}^k, \mathcal{B}_{tp}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tp}, \mathcal{B}_{tn}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tn}$ are abstract matrices for the system in \eqref{eq:observer_LTI_cf} and $r$ is a non-deterministic parametric input (\emph{ie} $\forall k>0, r_{k+1} =r_k$) whereas the set N is composed of non-deterministic variable inputs.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Controller synthesis}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
Our model, rather than look at a bisimulation \todo{specific notion that has not been formally define - do not use} in discrete time, explores a much higher abstraction, 
evaluating the unbounded time-continuous space reach tube as a set of inequalities that represent only the safety specification.

\begin{enumerate}
\item 
We start by preprocessing the dynamics. This can be computationally intensive and we don't want it to be part of the CEGIS loop.
\begin{enumerate}
\item Calculate $\mat{T}$ using equations \eqref{eq:rncf}\eqref{eq:rcf}\eqref{eq:to_cf}
\item Extract the Plant's characteristic polynomial coefficients from the last row of $P_a=(\mat{T}\mat{A}\mat{T}^-1)_{n*}$.
\item Calculate the noise set $N$ from the quantizer resolutions and estimated round-off errors ($\frac{q_1}{2},\frac{q_2}{2},q_3$)
\item Calculate a set of initial bounds on K based on the input constraints. $$\phi_{input} \Leftrightarrow u_k=r-Kx_k : x \in X^\# \wedge u_k \in [\underline{u}\ \ \overline{u}]$$
\end{enumerate}
\item Next we synthesise a candidate controller. 
We begin with a minimum set of linear inequalities describing the constraints caused by the input and the initial state during a single iteration, in addition to a stability specification defined by Jury's criteria. %, and try to synthesise a controller that meets these specs.

>>>>

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Verification of a controllable system}
\label{sec:cof_verification}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Whilst there are other properties, we may include in the verification, 
% we will solely focus on these for simplicity of the explanation.
% We assume that we do not want saturation in our system, 
% hence the control loop has only one mode.

We have also identified 
the quantization noise as a bounded non-deterministic input.
%
The first step constraints
\begin{equation}
\mat{K} \models u_k=-\mat{K}x_k : \phi_{safety}
\label{eq:of_bounds}
\end{equation}
The second step is evaluating the progression of the system over time to evaluate 
$x_{k+1} \models \phi_{safety}$. For this step we require 
an initial set from which the system progresses. Obviously, this set $x_0$ 
must satisfy the specification; otherwise, the system will be unsafe to begin with.
We accept a specification in the form $\mat{E}\mat{T}x_0<\mat{f}$. 
The presence of $\mat{T}$ is because this will typically be given in the original 
state-space. Accelerating \eqref{eq:observer_LTI_cf}, we obtain:
%
\begin{align}
\label{eq:acc_observer_LTI_cf}
x=&\mat{A}_t^kx_0
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tp} r_i
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tn}(\nu_1+\nu_2+\nu_3) : x_0 \models \phi_{init}
\end{align}
which would still require us to verify the system for every $k$ up to infinity. 
Instead we use abstract acceleration~\cite{cattaruzza2015unbounded}, which translates the formula into
%
\begin{align}
\label{eq:aa_observer_LTI_cf}
\hat{X}^\#
=&\mathcal{A}_t X_0+\mathcal{B}_{tp} R + \mathcal{B}_{tn} N\\
X_0 =&\left \{x : x \models \phi_{init} \right\}\nonumber\\
 R =&\{r : \mat{E}_r < f_r \}\nonumber\\
 N=&\left \{ \nu_1+\nu_2+ \nu_3 : \nu_1 \in \left[-\frac{q1}{2}\ \ \frac{q1}{2}\right] \wedge \nu_2 \in \left[-\frac{q2}{2}\ \ \frac{q2}{2}\right]  \wedge \nu_3 \in \left[-q3\ \ q3\right]  \right \}\nonumber
\end{align}
%
where $\mathcal{A}_t=\bigcup_{k=1}^\infty \mat{A}^k, \mathcal{B}_{tp}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tp}, \mathcal{B}_{tn}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tn}$ are abstract matrices for the system in \eqref{eq:observer_LTI_cf} and $r$ is a non-deterministic parametric input (\emph{ie} $\forall k>0, r_{k+1} =r_k$) whereas the set N is composed of non-deterministic variable inputs.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Controller synthesis}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
The CEGIS loop for this approach is illustrated in \ref{fig:CEGARIS}.

% Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
% Our model, rather than look at a bisimulation \todo{specific notion that has not been formally define - do not use} in discrete time, explores a much higher abstraction, 
% evaluating the unbounded time-continuous space reach tube as a set of inequalities that represent only the safety specification.

\begin{enumerate}
\item 
We start by preprocessing the dynamics. This can be computationally intensive and we don't want it to be part of the CEGIS loop.
% Let us recall the formulas used for single-input and single-output (SISO) systems 
% in sections \ref{sec:reachable} and \ref{sec:observable}.
A controllable system with quantization will have the following dynamics 
(note that we use underscript ``n'' to denote a noise term,
``cf'' to denote Controllable Canonical Form and $\mat{J}_{n,m}$ to represent a matrix of ones with $n$ rows and $m$ columns):
\todo{eliminate vectors}
%
\begin{align}
\label{eq:observer_LTI_cf}
x_{k+1}=&\mat{A}_{t}x_k+\mat{B}_{t} r_k+\mat{B}_{n} (\nu_1 + \nu_2 + \nu_3)
\end{align}
\begin{align*}
\mat{A}_{t}=&\mat{T}\mat{A}_d\mat{T}^{-1}&
\mat{B}_{t}=&\mat{T}\mat{B}_d&
\mat{B}_{n}=&\mat{J}_{n,1}&
\mat{T}=&\mat{R}_{cf}\mat{R}^{-1}&
\end{align*}

@@

\begin{enumerate}
\item Calculate $\mat{T}$ using equations \eqref{eq:rncf}\eqref{eq:rcf}\eqref{eq:to_cf}
\item Extract the Plant's characteristic polynomial coefficients from the last row of $P_a=(\mat{T}\mat{A}\mat{T}^-1)_{n*}$.
\item Calculate the noise set $N$ from the quantizer resolutions and estimated round-off errors ($\frac{q_1}{2},\frac{q_2}{2},q_3$)
\item Calculate a set of initial bounds on K based on the input constraints. $$\phi_{input} \Leftrightarrow u_k=r-Kx_k : x \in X^\# \wedge u_k \in [\underline{u}\ \ \overline{u}]$$
\end{enumerate}
\item Next we synthesise a candidate controller. 
We begin with a minimum set of linear inequalities describing the constraints caused by the input and the initial state during a single iteration, in addition to a stability specification defined by Jury's criteria. %, and try to synthesise a controller that meets these specs.
% For a stabilized closed-loop, which is necessary for our safety property, we need
% these roots to remain within the unit circle, which we verify using Jury's criteria.
% Notice that in our polynomials $c_0=1$ and we only use reals.
% Once we have established convergence, we may examine the 
% safety specification $\phi_{safety}$.
